---
title: "scFEA"
author: "MinJae Kim"
date: "2024-07-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

##### Load required libraries

```{r, message = FALSE}
library(Seurat)
library(ggplot2)
library(data.table)
library(stringr)
library(dplyr)
```

##### Set the data path for the scFEA analysis. Load required files.

```{r}
data_path <- '/Users/mj/Documents/lab/scFEA/hfd/'
hfd_obj <- readRDS(paste0(data_path, 'clean_hfd_harmony_unannotated_res0.55.rds'))
```

##### Create subdirectory named with today's date. Create directories named input and output within that folder.

```{r}
dir.create(file.path(data_path, Sys.Date()))
subdir_path <- paste0(data_path, Sys.Date())
input_dir <- paste0(file.path(subdir_path, "input"))
dir.create(input_dir)
output_dir <- paste0(file.path(subdir_path, "output"))
dir.create(output_dir)
```

##### Trim the original data to contain only genes used for the flux computation

```{r}
scFEA_mouse_genes <- read.table(paste0(data_path, '/scFEA.mouse.genes.txt'))
trimmed_hfd_obj <- subset(hfd_obj, features = scFEA_mouse_genes$V1)
```

##### Write the trimmed object into a .csv file

```{r}
write_out <- as.data.frame(as.matrix(trimmed_hfd_obj@assays$RNA@data))
fwrite(x = write_out, row.names = TRUE, file = paste0(input_dir, '/trimmed_hfd_data.csv'))
```

##### Go to http://scflux.org/, select the species and the network. Save the output files into /Users/mj/Documents/lab/scFEA/hfd/%date%/output

##### Load the web-generated output files (.csv)

```{r}
predFlux <- read.csv(paste0(output_dir, '/trimmed_hfd_data_flux.csv'), header = T, row.names = 1)
predFlux <- data.matrix(predFlux)
predFlux0 <- t(predFlux)
#If not using the web-based version, I found it necessary to substitute "." into "-" in the cell barcodes for the downstream analysis. May be a bug:
#colnames(predFlux0) <- gsub("\\.", "-", colnames(predFlux0))
```

##### Add Flux as a new assay

```{r}
hfd_obj@assays$FLUX <- CreateAssayObject(counts = predFlux0)
```

##### Change default assay to FLUX (can change back any time)

```{r}
DefaultAssay(hfd_obj) <- 'FLUX'
```

##### Create a list of names and annotate the clusters. Will make the graph easier to read
```{r}
moduleinfo <- read.csv(paste0(output_dir, '/scFEA.Glucose-TCAcycle.mouse.moduleinfo.csv'), header = T)
module.name <- moduleinfo['M_name']
module.id <- moduleinfo['M_id']
```

##### Create a list of modules. This will be useful for drawing plots

```{r}
nmodules <- ncol(predFlux)
modules <- colnames(predFlux)
modules <- gsub("\\_", "-", modules)
```

##### Rename/annotate cluster ids. This will make it easier to read plots
```{r}
cluster.ids <- c("ST2hi p1", "ST2hi p2", "Precursor-like", "TCF+ p1", "TCF+ p2", "TCF+ p3", "Tbet+", "ST2hi p3")
names(cluster.ids) <- levels(hfd_obj)
hfd_obj <- RenameIdents(hfd_obj, cluster.ids)
```



##### Show the original umap from the RNA data
```{r}
fig7d <- DimPlot(hfd_obj, reduction = "umap", split.by = "Groups") + ggtitle('Fig 7d')
fig7d
```

##### Find variable features, scale data
```{r}
hfd_obj <- FindVariableFeatures(hfd_obj, selection.method = "vst", nfeatures = 2000, verbose = F)
hfd_obj <- ScaleData(hfd_obj, features = rownames(hfd_obj), assay = 'FLUX', verbose = F)
```


##### Output RidgePlot
```{r}
for (i in 1:nmodules) {
  rplt <- (RidgePlot(hfd_obj, features = modules[i], group.by = "Groups") & ggtitle(module.name[i,]))
  print(rplt)
}
```
##### Output Dotplot and VlnPlot

```{r}
for (i in 1:nmodules) {
  vplt <- (VlnPlot(hfd_obj, pt.size = 0.3, features = modules[i], group.by = "Groups") & ggtitle(module.name[i,]))
  print(vplt)
}
```

```{r}
ae <- AverageExpression(hfd_obj, assays = 'FLUX', layer = 'data', group.by = 'Groups')
ae <- as.data.frame(as.matrix(ae))
ae <- as.data.frame(as.matrix(ae$V1$FLUX))
ae$modules <- module.name$M_name
ae

DotPlot(hfd_obj, features = modules, group.by = 'Groups', scale = FALSE) + RotatedAxis()
DotPlot(hfd_obj, features = modules, split.by = "Groups") + RotatedAxis()
DotPlot(hfd_obj, features = modules, split.by = "Groups") + RotatedAxis()
DotPlot(hfd_obj, features = modules[0:(nmodules/2)], split.by = "Groups") + RotatedAxis()
DotPlot(hfd_obj, features = modules[(nmodules/2):nmodules], split.by = "Groups") + RotatedAxis()
```

##### Output Heatmap
```{r}
DoHeatmap(hfd_obj, features = modules[0:(nmodules/2)], group.by = "Groups")
DoHeatmap(hfd_obj, features = modules[(nmodules/2):nmodules], group.by = "Groups")
```

##### Output FeaturePlot
```{r}
for (i in 1:nmodules) {
  fplot <- FeaturePlot(hfd_obj, pt.size = 2, features = modules[i], split.by = "Groups") & ggtitle(module.name[i,])
  print(fplot)
}
```

